### 2018-06-25 reports/analysis_pilot.Rmd
add analysis per each category
currently on one subject only

---
title: "Analysis"
author: "Filip Dechterenko"
date: "25 června 2018"
output: html_document
---

# Analysis categ

## Load data

```{r}

library(tidyverse)

data_pth <- here::here("data", "results", "categ")

col_types <- cols(
  .default = col_character(),
  prot_id = col_integer(),
  trial_id = col_integer(),
  target_position = col_integer(),
  quintile = col_integer(),
  trials.thisRepN = col_integer(),
  trials.thisTrialN = col_integer(),
  trials.thisN = col_integer(),
  trials.thisIndex = col_integer(),
  mouse.x = col_double(),
  mouse.y = col_double(),
  mouse.leftButton = col_integer(),
  mouse.midButton = col_integer(),
  mouse.rightButton = col_integer(),
  protocol_id = col_integer(),
  subject_id = col_integer(),
  frameRate = col_double()
)

df <- data_pth %>% 
  dir(full.names = T) %>% 
  purrr::map(read_csv, col_types = col_types) %>% 
  bind_rows()

df <- df %>%
  mutate(target_image = paste0("image",target_position),
         correct = as.numeric(target_image == mouse.clicked_name),
         category_type = if_else(category %in% c("desert", "waves", "coast", "underwater", "iceberg","field","garden","cavern","beach","canyon","mountainwhite","woods"),"natural","manmade"))

```

# Desc stat

```{r}
df %>% 
  group_by(quintile) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2))

df %>% 
  group_by(category_type) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2))

```

# Visualize

```{r}
df %>% ggplot(aes(x = quintile, y = correct)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) 
```

# Analysis all

## Load data

```{r}

library(tidyverse)

data_pth <- here::here("data", "results", "all")

col_types <- cols(
  .default = col_character(),
  prot_id = col_integer(),
  trial_id = col_integer(),
  target_position = col_integer(),
  quintile = col_integer(),
  trials.thisRepN = col_integer(),
  trials.thisTrialN = col_integer(),
  trials.thisN = col_integer(),
  trials.thisIndex = col_integer(),
  mouse.x = col_double(),
  mouse.y = col_double(),
  mouse.leftButton = col_integer(),
  mouse.midButton = col_integer(),
  mouse.rightButton = col_integer(),
  protocol_id = col_integer(),
  subject_id = col_integer(),
  frameRate = col_double()
)

df <- data_pth %>% 
  dir(full.names = T) %>% 
  purrr::map(read_csv, col_types = col_types) %>% 
  bind_rows()

df <- df %>%
  mutate(target_image = paste0("image",target_position),
         correct = as.numeric(target_image == mouse.clicked_name))

```

```{r}
df %>% ggplot(aes(x = quintile, y = correct)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  ylim(0,1)+
  theme(aspect.ratio = 1)
```