### 2018-06-28 reports/analysis_pilot.Rmd
Finished testing of pilot data

---
title: "Analysis"
author: "Filip Dechterenko"
date: "25 června 2018"
output: html_document
---

# Analysis categ

## Load data

```{r, warning=FALSE}

load(here::here("data","konkle_180621.RData"))

library(tidyverse)

data_pth <- here::here("data", "results", "categ")

col_types <- cols(
  .default = col_character(),
  prot_id = col_integer(),
  trial_id = col_integer(),
  target_position = col_integer(),
  quintile = col_integer(),
  trials.thisRepN = col_integer(),
  trials.thisTrialN = col_integer(),
  trials.thisN = col_integer(),
  trials.thisIndex = col_integer(),
  mouse.x = col_double(),
  mouse.y = col_double(),
  mouse.leftButton = col_integer(),
  mouse.midButton = col_integer(),
  mouse.rightButton = col_integer(),
  protocol_id = col_integer(),
  subject_id = col_integer(),
  frameRate = col_double()
)

df <- data_pth %>% 
  dir(pattern = "*.csv", full.names = T) %>% 
  purrr::map(read_csv, col_types = col_types) %>% 
  bind_rows()

df <- df %>%
  mutate(target_image = paste0("image",target_position),
         correct = as.numeric(target_image == mouse.clicked_name),
         category_type = if_else(category %in% c("desert", "waves", "coast", "underwater", "iceberg","field","garden","cavern","beach","canyon","mountainwhite","woods"),"natural","manmade"))

```

# Desc stat

```{r}
df %>% 
  group_by(quintile) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2))

df %>% 
  group_by(category_type) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2))

df %>% 
  group_by(category) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2), n= n()) %>% 
  arrange(-m)

df %>% 
  group_by(category,quintile) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2), n= n()) %>% 
  arrange(category,-m)

```

# Quintiles

## Visualize

```{r}
df %>% ggplot(aes(x = quintile, y = correct)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9)

df %>% ggplot(aes(x = quintile, y = correct)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  facet_wrap(~subject_id) +
  geom_hline(yintercept = 1/9) 
```

## Test

```{r lmer}
library(lme4)

glm1 <- glmer(cbind(correct,1-correct) ~ quintile + (1|subject_id) + (1|category), df, family = binomial())
glm1_null <- glmer(cbind(correct,1-correct) ~ (1|subject_id) + (1|category), df, family = binomial())
anova(glm1, glm_null)

```

# Express the distances as ratio

We are now sampling from the quintiles, we express the distances as ratio of average within group distance and between group distance

```{r}
n <- nrow(df)
fc7_names <- colnames(fc7)
df$d_ratio <- NA_real_ 
for (i in 1:n) {
  target_image <- df[[paste0("im",df$target_position[i])]][i]
  center_image <- df$selected_image[i]
  rest_images  <- df[i,paste0("im", setdiff(1:9,df$target_position[i]))] %>% as.matrix() %>% c()
  d_rest <- fc7[center_image == fc7_names, fc7_names %in% rest_images]
  df$d_ratio[i] <- fc7[center_image == fc7_names, target_image == fc7_names] / mean(d_rest[d_rest>0])

}

```

## Visualize

```{r}
df %>% ggplot(aes(x = d_ratio, y = correct)) + 
  geom_point() +
  stat_smooth(method="glm", method.args = list(family = "binomial"), se=F) +
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9)

df %>% ggplot(aes(x = d_ratio, y = correct)) + 
  geom_point() +
  stat_smooth(method="glm", method.args = list(family = "binomial"), se=F) +
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9)

```

## Test

```{r lmer}
library(lme4)

glm2 <- glmer(cbind(correct,1-correct) ~ d_ratio + (1|subject_id) + (1|category), df, family = binomial())
glm_null <- glmer(cbind(correct,1-correct) ~ (1|subject_id) + (1|category), df, family = binomial())
anova(glm2, glm_null)

```

# Correlation between raters

```{r ICC}
x <- df %>% select(prot_id, subject_id, correct) %>%
  group_by(prot_id) %>%
  summarize(ICC = psych::ICC(cbind()))
cor(df$correct[df$subject_id == 1],df$correct[df$subject_id == 7])
cor(df$correct[df$subject_id == 8],df$correct[df$subject_id == 100])
cor(df$correct[df$subject_id == 3],df$correct[df$subject_id == 9])
cor(df$correct[df$subject_id == 4],df$correct[df$subject_id == 10])
cor(df$correct[df$subject_id == 5],df$correct[df$subject_id == 11])
cor(df$correct[df$subject_id == 6],df$correct[df$subject_id == 12])

```