### 2019-04-25 reports/DRM_exp/analysis_DRM_exp1.Rmd
start of analysis of DRM exp

---
title: "Analysis"
author: "Filip Dechterenko"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

```{r}
set.seed(190425)
library(tidyverse)
theme_set(theme_bw())
library(lme4)
```

# Load data

```{r load data, warning=FALSE, message=FALSE}

df <- readRDS(here::here("data/exp_DRM1/results_190425.rds"))

plots_dir <- here::here("plots")

if(!dir.exists(plots_dir)) {
  dir.create(plots_dir)
}

df_acc <- df %>% filter(!is.na(correct))
```

# Desc stat

```{r desc stat}
df_acc %>% 
  group_by(type) %>% 
  summarize(m = mean(correct), sd = sd(correct)) %>% 
  knitr::kable(digits = 2)

df_acc %>% 
  group_by(subject_id, type) %>% 
  summarize(correct = mean(correct)) %>% 
  ggplot(aes(x = type, y = correct)) +
  stat_summary(fun.data = "mean_cl_boot") +
  ylim(0,1) + theme(aspect.ratio = 1)

df_acc %>% 
  group_by(subject_id, type) %>% 
  summarize(correct = mean(correct)) %>% 
  ggplot(aes(x = type, y = correct)) +
  geom_point() +
  ylim(0,1) + theme(aspect.ratio = 1) +
  facet_grid(~subject_id)

df %>% group_by(subject_id) %>% 
  summarize(correct = mean(correct),
            zH = qno)

df %>% 
  group_by(category) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2), n= n()) %>% 
  arrange(-m)

df %>% 
  group_by(category,quintile) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2), n= n()) %>% 
  arrange(category,-m)

```

# Quintiles

## Visualize

```{r}
p <- df %>% ggplot(aes(x = quintile, y = correct)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +
  theme(text = element_text(size=18)) +
  scale_x_continuous("Quintile", breaks = c(2,3,4), labels = c("2","3","4"))
p
ggsave(file.path(plots_dir, "opam_figure.eps"), p, width = 6, height = 6)

df %>% ggplot(aes(x = quintile, y = correct)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  facet_wrap(~subject_id) +
  geom_hline(yintercept = 1/9) 
```

## Test

```{r lmer - quintiles}
library(lme4)

glm1 <- glmer(cbind(correct,1-correct) ~ quintile + (1|subject_id) + (1|category), df, family = binomial())
glm1_null <- glmer(cbind(correct,1-correct) ~ (1|subject_id) + (1|category), df, family = binomial())
anova(glm1, glm1_null)

```

# Express the distances as ratio

We are now sampling from the quintiles, we express the distances as ratio of average within group distance and between group distance

```{r}
n <- nrow(df)
fc7_names <- colnames(fc7)
df$d_ratio <- NA_real_ 
for (i in 1:n) {
  target_image <- df[[paste0("im",df$target_position[i])]][i]
  center_image <- df$selected_image[i]
  rest_images  <- df[i,paste0("im", setdiff(1:9,df$target_position[i]))] %>% as.matrix() %>% c()
  d_rest <- fc7[center_image == fc7_names, fc7_names %in% rest_images]
  df$d_ratio[i] <- fc7[center_image == fc7_names, target_image == fc7_names] / mean(d_rest[d_rest>0])

}

```

## Visualize

```{r}
df %>% ggplot(aes(x = d_ratio, y = correct)) + 
  geom_point() +
  stat_smooth(method="glm", method.args = list(family = "binomial"), se=F) +
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9)

df %>% ggplot(aes(x = d_ratio, y = correct)) + 
  geom_point() +
  stat_smooth(method="glm", method.args = list(family = "binomial"), se=F) +
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  facet_wrap(~subject_id) +
  geom_hline(yintercept = 1/9)

```

## Test

```{r lmer d_ratio}

glm2 <- glmer(cbind(correct,1-correct) ~ d_ratio + (1|subject_id) + (1|category), df, family = binomial())
glm2_null <- glmer(cbind(correct,1-correct) ~ (1|subject_id) + (1|category), df, family = binomial())
anova(glm2, glm2_null)

```

# Correlation between raters

```{r ICC}
cor(df$correct[df$subject_id == 1],df$correct[df$subject_id == 7])
cor(df$correct[df$subject_id == 8],df$correct[df$subject_id == 100])
cor(df$correct[df$subject_id == 3],df$correct[df$subject_id == 9])
cor(df$correct[df$subject_id == 4],df$correct[df$subject_id == 10])
cor(df$correct[df$subject_id == 5],df$correct[df$subject_id == 11])
cor(df$correct[df$subject_id == 6],df$correct[df$subject_id == 12])

```