### 2019-04-26 reports/DRM_exp/analysis_DRM_exp1.Rmd
add classical dprime analysis

---
title: "Analysis"
author: "Filip Dechterenko"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

```{r}
set.seed(190425)
library(tidyverse)
theme_set(theme_bw())
library(lme4)
```

# Load data

```{r load data, warning=FALSE, message=FALSE}

df <- readRDS(here::here("data/exp_DRM1/results_190425.rds"))

plots_dir <- here::here("plots")

if(!dir.exists(plots_dir)) {
  dir.create(plots_dir)
}
```

```{r filter participant which was not paying attention}
df <- df %>% filter(subject_id != 5)
```

# Desc stat

```{r desc stat}
df %>% 
  group_by(type) %>% 
  summarize(m = mean(correct), sd = sd(correct)) %>% 
  knitr::kable(digits = 2)

df %>% 
  group_by(subject_id, type) %>% 
  summarize(correct = mean(correct)) %>% 
  ggplot(aes(x = type, y = correct)) +
  stat_summary(fun.data = "mean_cl_boot") +
  ylim(0,1) + theme(aspect.ratio = 1)

df %>% 
  group_by(subject_id, type) %>% 
  summarize(correct = mean(correct)) %>% 
  ggplot(aes(x = type, y = correct)) +
  geom_point() +
  ylim(0,1) + theme(aspect.ratio = 1) +
  facet_grid(~subject_id)

df_dprime <- df %>% group_by(subject_id,type) %>% 
  summarize(correct = mean(correct),
            zH = qnorm(correct)) %>% 
  mutate(d_old_close = zH[type == "target"]-zH[type == "close distractor"],
         d_old_far = zH[type == "target"]-zH[type == "far distractor"],
         d_close_far = zH[type == "close distractor"]-zH[type == "far distractor"])

df_dprime_agg <- df_dprime %>% 
  group_by(type) %>% 
  summarize(m = mean(zH),sd = sd(zH))

df_dprime_agg

x0 <- (-400:400)/100

y_close  <- dnorm(x0, df_dprime_agg$m[df_dprime_agg$type == "close distractor"], 1)
y_far  <- dnorm(x0, df_dprime_agg$m[df_dprime_agg$type == "far distractor"], 1)
y_target  <- dnorm(x0, df_dprime_agg$m[df_dprime_agg$type == "target"], 1)

df_plot <- tibble(x = rep(x0,3), y = c(y_close, y_far, y_target), type = rep(c("close","far","target"), each = length(x0)))

p_dprimes <- df_plot %>% ggplot(aes(x = x, y = y, col = type)) + geom_point() + geom_vline(xintercept = 0) +
  theme(aspect.ratio = 1)
p_dprimes

```

split the analysis for first half and second half.

```{r}
trial_ids <- df %>% 
  filter(subject_id == 1) %>% 
  pull(trial_id)

cutoff_trial <- trial_ids[round(length(trial_ids)/15/2)*15]

df_acc2 <- df %>% mutate(part = if_else(trial_id<=cutoff_trial, "part1", "part2"))

df_acc2 %>% 
  group_by(type,part) %>% 
  summarize(m = mean(correct), sd = sd(correct)) %>% 
  knitr::kable(digits = 2)

df_acc2 %>% 
  group_by(subject_id, type, part) %>% 
  summarize(correct = mean(correct)) %>% 
  ggplot(aes(x = type, y = correct, col = part, group = part)) +
  stat_summary(fun.data = "mean_cl_boot") +
  stat_summary(fun.y = "mean", geom = "line") +
  ylim(0,1) + theme(aspect.ratio = 1)

```

```{r two dprimes and bias}
df_dprime2 <- df %>% 
  group_by(subject_id) %>% 
  summarize(H = sum(key_resp == "old" & corrKey == "old")/(sum(corrKey == "old")),
            FA_close = sum(key_resp == "old" & corrKey == "new" & type == "close distractor")/(sum(corrKey == "new" & type == "close distractor")),
            FA_far = sum(key_resp == "old" & corrKey == "new" & type == "far distractor")/(sum(corrKey == "new" & type == "far distractor"))) %>% 
  mutate(d_oldclose = qnorm(H) - qnorm(FA_close),
         d_oldfar = qnorm(H) - qnorm(FA_far),
         c_oldclose = -0.5*(qnorm(H) + qnorm(FA_close)),
         c_oldfar = -0.5*(qnorm(H) + qnorm(FA_far)))
t.test(df_dprime2$d_oldclose, df_dprime2$d_oldfar)
t.test(df_dprime2$c_oldclose, df_dprime2$c_oldfar)  
colMeans(df_dprime2)
```

```{r use rating to compute ROC}

df2 <- df %>% unite(resp, key_resp, confidence) %>% filter(type!="far distractor")
df3 <- df %>% unite(resp, key_resp, confidence) %>% filter(type!="close distractor")

tb1_oldclose <- table(df2$corrKey,df2$resp) %>% prop.table(margin = 1)
tb1_oldclose <- tb1_oldclose[c(2,1),c(6,4,5,1,2,3)]
tb1_oldclose[1,] <- cumsum(tb1_oldclose[1,])
tb1_oldclose[2,] <- cumsum(tb1_oldclose[2,])
tb1_oldclose <- tb1_oldclose[,-6]

tb2_oldclose <- qnorm(tb1_oldclose)
dprimes_oldclose <- tb2_oldclose[1,]-tb2_oldclose[2,]

tb1_oldfar <- table(df3$corrKey,df3$resp) %>% prop.table(margin = 1)
tb1_oldfar <- tb1_oldfar[c(2,1),c(6,4,5,1,2,3)]
tb1_oldfar[1,] <- cumsum(tb1_oldfar[1,])
tb1_oldfar[2,] <- cumsum(tb1_oldfar[2,])
tb1_oldfar <- tb1_oldfar[,-6]

tb2_oldfar <- qnorm(tb1_oldfar)
dprimes_oldfar <- tb2_oldfar[1,]-tb2_oldfar[2,]

HFA  <- rbind(tibble(H =tb1_oldclose[1,], FA = tb1_oldclose[2,], type = "old-close"),
              tibble(H =tb1_oldfar[1,], FA = tb1_oldfar[2,], type = "old-far"))
zHFA <- rbind(tibble(H =tb2_oldclose[1,], FA = tb2_oldclose[2,], type = "old-close"),
              tibble(H =tb2_oldfar[1,], FA = tb2_oldfar[2,], type = "old-far"))
HFA %>% ggplot(aes(x = FA, y = H, col = type)) + 
  geom_point() + 
  xlim(0,1) + 
  ylim(0,1) + 
  geom_segment(x = 0, y = 0, xend = 1,yend = 1) + theme(aspect.ratio = 1)

zHFA %>% ggplot(aes(x = FA, y = H)) + 
  geom_point() + 
  stat_smooth(method = "lm", se = F) + 
  xlim(-2.5,2.5) + 
  ylim(-2.5,2.5) + 
  geom_segment(x = -2, y = -2, xend = 2,yend = 2)
```

# Memorability

Memorability were computed using pretrained CNN

```{r load mem data}
mem_imgnames <- read_csv(here::here("data/exp_DRM1/memorability/memorability_imgnames.txt"), col_names = F)
mem_values <- read_csv(here::here("data/exp_DRM1/memorability/memorability_all.txt"), col_names = F)

df_mem <- tibble(img_name = mem_imgnames$X1, mem_score = mem_values$X1)
dfm <- df %>% left_join(df_mem,by = "img_name")

cor.test(~correct+mem_score, dfm %>% filter(type == "target"))
cor.test(~correct+mem_score, dfm %>% filter(type == "close distractor"))
cor.test(~correct+mem_score, dfm %>% filter(type == "far distractor"))
```

# Per image accuracy

```{r }
df %>% group_by(img_name,type) %>% summarize(m = mean(correct), sd = sd(correct)) %>% arrange(-m)
```