### 2019-06-27 reports/odd_one_out/analysis_oddoneout.Rmd
change some translations

---
title: "Analysis of Odd one out experiment"
author: "Filip Dechterenko"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

# Load libraries

```{r load libraries, warning=FALSE,message=FALSE}
set.seed(167)
library(tidyverse); theme_set(theme_classic(24))
library(lme4)
library(MuMIn)
library(here)

source(here("R","utils.R"))
```

# Analysis categ

First analyze data when computations were done wihtin a category

## Load data

```{r load data, warning=FALSE, message=FALSE}

load(here("data","konkle_180621.RData"))

df <- readRDS(here("data","oddoneout","results_190531.rds"))

metrics <- readRDS(file = here("data", "oddoneout", "metrics_190614.rds"))

wrong_trials <- df %>% select(prot_id_version,trial_id,im1:im9) %>% group_by(prot_id_version,trial_id) %>% gather("key", "value",im1:im9) %>% summarize(n = length(unique(value))) %>% filter(n<9) %>% select(-n)

df_withoutwrong <- df %>% anti_join(wrong_trials, by = c("prot_id_version","trial_id"))

participants <- readxl::read_excel(here("data","oddoneout","oddoneout_participants.xlsx"))

```

```{r create plot dir}
plots_dir <- here("plots", "oddoneout")

if(!dir.exists(plots_dir)) {
  dir.create(plots_dir, recursive = T)
}

```

# Describe the participants

```{r describe participants}
participants %>% 
  summarize(m_age = mean(age),sd_age = sd(age), n_males = sum(gender == "M")) %>% 
  knitr::kable(digits = 2)
```

# Duration of the exp

```{r duration of the exp}
data_pth <- here("data", "oddoneout", "results_categ")

exp_duration <- data_pth %>% 
  dir(pattern = "*.log", full.names = T) %>% 
  purrr::map(read_file) %>% 
  purrr::map_dbl(get_dur)

```

Average experiment duration: `r (exp_duration / 60) %>% mean() %>% round(2)`, SD = `r (exp_duration / 60) %>% sd() %>% round(2)`

# Desc stat

```{r desc stat}
df %>% 
  group_by(subject_id, quintile) %>% 
  summarize(correct = mean(correct)) %>% 
  group_by(quintile) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2))

df %>% 
  group_by(category_type) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2))

df %>% 
  group_by(category) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2), n= n()) %>% 
  arrange(-m)

df %>% 
  group_by(category,quintile) %>% 
  summarize(m = mean(correct) %>% round(2), sd = sd(correct) %>% round(2), n= n()) %>% 
  arrange(category,-m)

```

# Quintiles

## Histogram

```{r}

this_cat <- df %>% filter(subject_id == 1, trial_id == 1) %>% pull(category)
this_center <- df %>% filter(subject_id == 1, trial_id == 1) %>% pull(selected_image)

df_onecat_distances <- fc7[fc7_names == this_center, str_starts(fc7_names, this_cat)] %>% enframe() %>% rename(distance = value)

qs <- compute_quintiles(df_onecat_distances$distance) 
df_onecat_distances <- df_onecat_distances %>% 
  mutate(quintile = case_when(
    distance<qs[2] ~ 1,
    distance >= qs[2] & distance < qs[3] ~ 2,
    distance >= qs[3] & distance < qs[4] ~ 3,
    distance >= qs[4] & distance < qs[5] ~ 4,
    TRUE ~ 5
  ) %>% as.factor())

df_segments <- tibble(x = qs[2:6], xend = x, y = 0, yend = 17)

p <- df_onecat_distances %>% 
  ggplot(aes(x = distance)) + 
  geom_histogram(bins = 30) + 
  theme(aspect.ratio = 1) +
  geom_segment(data = df_segments, aes(x=x,xend=xend,y=y,yend=yend)) +
  annotate(x = qs[3:5]-c(3,1,-0.6), y = c(19,19,19), label = c("2.","3.","4."), geom = "label", size = 5) +
  ggtitle(sprintf("Category: %s",this_cat)) +
  xlab("Distance") +
  ylab("Count")
p  
ggsave(file.path(plots_dir, "Fig_chp3_distance_histogram.svg"), p, width = 6, height = 6)  
```

## Visualize

```{r}
p <- df %>%
  group_by(subject_id, quintile) %>% 
  summarize(correct = mean(correct)) %>% 
  mutate(quintile = as.factor(quintile)) %>% 
  ggplot(aes(x = quintile, y = correct)) + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +

  #scale_x_continuous("Kvintil", breaks = c(2,3,4), labels = c("2","3","4")) + 
  geom_violin(trim = F)+
  stat_summary(fun.data = "mean_cl_boot") + 
  xlab("Quintile")

p
ggsave(file.path(plots_dir, "Fig_chp3_results.eps"), p, width = 6, height = 6)

p <- df %>%
  mutate(quintile = as.factor(quintile)) %>% 
  ggplot(aes(x = quintile, y = correct)) + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +

  #scale_x_continuous("Kvintil", breaks = c(2,3,4), labels = c("2","3","4")) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  xlab("Quintile") + 
  facet_wrap(~subject_id, ncol = 4)

p
ggsave(file.path(plots_dir, "Fig_chp3_results_participants.eps"), p, width = 6, height = 6)

```

## Visualize categories

```{r}
order_category <- 
  df %>%
  group_by(category,category_type) %>% 
  do(Hmisc::smean.cl.boot(.$correct) %>% t() %>% as_tibble()) %>% 
  arrange(Mean) %>% 
  ungroup() %>% 
  mutate(category = forcats::as_factor(category), rank = 1:n())

p_ordercat <- order_category %>%
  ggplot(aes(x = category, xend = category, y = Lower, yend = Upper, group = 1, col = category_type)) +
  geom_segment(size = 1.2) +
  geom_point(aes(y=Mean), size = 3) + 
  theme(aspect.ratio = 0.2) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size = 16)) + 
  theme(legend.position = "top") + 
  xlab("Category") + 
  ylab("Perc. correct") + 
  scale_color_discrete("Category type", labels = c("Man-made","Natural")) 
p_ordercat
ggsave(file.path(plots_dir, "Fig_chp3_ordercat.svg"), p_ordercat, width = 18, height = 6)
```

## Test

```{r lmer - quintiles}
glm1 <- glmer(correct ~ quintile + (1|subject_id) + (1|category), df, family = binomial())
glm1_null <- glmer(correct ~ (1|subject_id) + (1|category), df, family = binomial())
anova(glm1, glm1_null)

```

```{r}
df_for_onettest <- df %>% group_by(subject_id,quintile) %>% summarize(correct = mean(correct))
df_for_onettest %>% group_by(quintile) %>% do(t.test(.$correct, mu = 1/9) %>% broom::tidy()) %>% 
  left_join(df_for_onettest %>% group_by(quintile) %>% do(d=lsr::cohensD(.$correct, mu = 1/9)) %>% mutate(d = unlist(d)),
            by = "quintile")

```
# Express the distances as ratio

We are now sampling from the quintiles, we express the distances as ratio of average within group distance and between group distance

```{r}
n <- nrow(df)
fc7_names <- colnames(fc7)
df$d_ratio <- NA_real_ 
for (i in 1:n) {
  target_image <- df[[paste0("im",df$target_position[i])]][i]
  center_image <- df$selected_image[i]
  rest_images  <- df[i,paste0("im", setdiff(1:9,df$target_position[i]))] %>% as.matrix() %>% c()
  d_rest <- fc7[center_image == fc7_names, fc7_names %in% rest_images]
  df$d_ratio[i] <- fc7[center_image == fc7_names, target_image == fc7_names] / mean(d_rest[d_rest>0])

}

```

## Test

```{r lmer d_ratio}

glm2 <- glmer(correct ~ d_ratio + (1|subject_id) + (1|category), df, family = binomial(),  na.action = "na.fail")
glm2_null <- glmer(correct ~ (1|subject_id) + (1|category), df, family = binomial())
anova(glm2, glm2_null)

```

### Visualize glmer

```{r}
#dredge.models<-dredge(glm2,trace=FALSE,rank="AICc")
#my.dredge.models<-get.models(dredge.models,subset = T)
#silly<-model.avg(my.dredge.models,subset=delta<10)
#predict(silly,type="response")

p0 <- df %>% ggplot(aes(x = d_ratio, y =correct))+
    geom_point(alpha=0.5)

predframe <- tibble(d_ratio= df$d_ratio)

predframe$correct <- predict(glm2,newdata=predframe,type="response",re.form=NA)
p <- p0 + geom_point(data=predframe,colour="red")+
  geom_line(data=predframe,colour="red",aes(group=1)) + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  xlab("Distance ratio") + 
  ylab("Perc. correct")

p
ggsave(file.path(plots_dir, "Fig_chp3_results_ratio.svg"), p, width = 6, height = 6)

```

## Visualize top and bottom category

```{r top and bottom}
set.seed(190610)
onecat_plot <- function(cat_name) {
  df_thiscat <- df %>% filter(category == cat_name)

  img_names <- df_thiscat %>% select(im1:im9) %>% gather() %>% pull(value) %>% unique()  
  mds_df <- cmdscale(fc7[fc7_names %in% img_names, fc7_names %in% img_names]) %>% as.data.frame() %>% rownames_to_column()

  df1 <- df_thiscat %>% filter(quintile == 4) %>% sample_n(1)

  target_image <- df1[[paste0("im",df1$target_position)]]   
  center_image <- df1$selected_image
  rest_images  <- df1[,paste0("im", setdiff(1:9,df1$target_position))] %>% as.matrix() %>% c()

  df_layerdata <- df1 %>% select(im1:im9) %>% gather() %>% rename(img_name = value)
  df_layerdata <- df_layerdata %>% 
    mutate(type = if_else(img_name == target_image,"target",
                          if_else(img_name == center_image, "center", "distractors"))) %>% left_join(mds_df, by = c("img_name"="rowname"))

  p <- ggplot(mds_df, aes(x = V1, y = V2), size = 2) + geom_point() +
    geom_point(data = df_layerdata, aes(x = V1, y = V2, col = type), size = 2) +

    theme(aspect.ratio = 1) + 
    xlab("MDS 1") + ylab("MDS 2") + 
    ggtitle(sprintf("Category: %s", cat_name)) +
    scale_color_manual(values = c("green","blue","red"))
  p
}
bottom_one <- order_category %>% filter(rank == 1)
top_one <- order_category %>% filter(rank == 64)
fc7_names <- colnames(fc7)

p1 <- onecat_plot(bottom_one$category[1])
p2 <- onecat_plot(top_one$category[1])
ggsave(file.path(plots_dir, "Fig_chp3_category_closet.svg"), p1, width = 6, height = 6)
ggsave(file.path(plots_dir, "Fig_chp3_category_field.svg"), p2, width = 6, height = 6)
```

# Aggreement between raters

```{r ICC}
ck1 <- psych::cohen.kappa(cbind(df$mouse.clicked_name[df$subject_id == 1],df$mouse.clicked_name[df$subject_id == 7]))
ck2 <- psych::cohen.kappa(cbind(df$mouse.clicked_name[df$subject_id == 8],df$mouse.clicked_name[df$subject_id == 100]))
ck3 <- psych::cohen.kappa(cbind(df$mouse.clicked_name[df$subject_id == 3],df$mouse.clicked_name[df$subject_id == 9]))
ck4 <- psych::cohen.kappa(cbind(df$mouse.clicked_name[df$subject_id == 4],df$mouse.clicked_name[df$subject_id == 10]))
ck5 <- psych::cohen.kappa(cbind(df$mouse.clicked_name[df$subject_id == 5],df$mouse.clicked_name[df$subject_id == 11]))
ck6 <- psych::cohen.kappa(cbind(df$mouse.clicked_name[df$subject_id == 6],df$mouse.clicked_name[df$subject_id == 12]))

cks <- list(ck1,ck2,ck3,ck4,ck5,ck6)
cks <- cks %>% purrr::map_dbl(~.$weighted.kappa)
```

Average agreement between raters was `r mean(cks) %>% round(2)

## Metrics

```{r}
p <- metrics %>% 
  filter(metric!= "ssim_correct" & metric!= "correct") %>% 
  ggplot(aes(x = quintile, y = value_fc7, col = metric, group = metric)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +
  scale_x_continuous("Quintile", breaks = c(2,3,4), labels = c("2","3","4")) +
  theme(text = element_text(size=18)) + scale_color_discrete(labels = c("GIST", "HOG", "RGB", "SIFT")) + 
  stat_summary(fun.y=mean, geom="line") +
  ggtitle("Predict CNN")
  #theme(legend.justification = c(1,1), legend.position=c(0.99,0.99))
p
ggsave(file.path(plots_dir, "Fig_chp3_metrics_explanation_fc7.svg"), p, width = 6, height = 6)

p <- metrics %>% 
  filter(metric!= "ssim_correct") %>% 
  ggplot(aes(x = quintile, y = value_partic, col = metric, group = metric)) + 
  stat_summary(fun.data = "mean_cl_boot") + 
  ylim(0,1)+
  theme(aspect.ratio = 1) +
  geom_hline(yintercept = 1/9) +
  ylab("Perc. correct") +
  scale_x_continuous("Quintile", breaks = c(2,3,4), labels = c("2","3","4")) +
  theme(text = element_text(size=18)) + scale_color_discrete(labels = c("Humans", "GIST", "HOG", "RGB", "SIFT")) + 
  stat_summary(fun.y=mean, geom="line") +
  ggtitle("Predict human peformance")
  #theme(legend.justification = c(1,1), legend.position=c(0.99,0.99))
p
ggsave(file.path(plots_dir, "Fig_chp3_metrics_predicthuman.svg"), p, width = 6, height = 6)

```