### 2020-03-05 reports/DRM_exp/analysis_DRM_after_review.Rmd
new analysis

---
title: "Analysis after review"
author: "Filip Dechterenko"
date: '`r format(Sys.time(), "%d %B, %Y")`'
output: html_document
---

# Load libraries

```{r libraries and preparation}
set.seed(190425)
library(tidyverse); theme_set(theme_classic(16))
library(lme4)
library(here)

source(here("R","utils.R"))
```

# Load data

Data were preprocesseb by scripts

```{r load data, warning=FALSE, message=FALSE}

df <- readRDS(here("data", "exp_DRM1", "results_191017.rds"))

df_participants <- readxl::read_excel(here::here("data","exp_DRM1","participants_190509.xlsx"))

plots_dir <- here("plots", "exp_DRM1")

if(!dir.exists(plots_dir)) { dir.create(plots_dir) }

df <- df %>% left_join(df_participants, by = c("subject_id"))

df2 <- readRDS(here("data/exp_DRM1/results_with_tgtdistances_fc7_200219.rds"))
df2_conv5 <- readRDS(here("data/exp_DRM1/results_with_tgtdistances_conv5_200219.rds"))
df2_conv3 <- readRDS(here("data/exp_DRM1/results_with_tgtdistances_conv3_200219.rds"))
```

## Correlation of distance from prototype with accuracy

```{r Correlation of distance from prototype with accuracy }

cor.test(~dist_to_proto+correct, df2)
cor.test(~dist_to_proto+correct, df2 %>% filter(type == "far distractor"))
cor.test(~dist_to_proto+correct, df2 %>% filter(type == "close distractor"))
cor.test(~dist_to_proto+correct, df2 %>% filter(type == "target"))

cor.test(~dist_to_proto_conv5+correct, df2_conv5)
cor.test(~dist_to_proto_conv5+correct, df2_conv5 %>% filter(type == "far distractor"))
cor.test(~dist_to_proto_conv5+correct, df2_conv5 %>% filter(type == "close distractor"))
cor.test(~dist_to_proto_conv5+correct, df2_conv5 %>% filter(type == "target"))

```

### Distances from tgt images - fc7

#### fc7

```{r}
gt0 <- function(x) {x >0}

df2_fc7 <- df2 %>% 
  rowwise() %>% 
  mutate(mean_tgt_dist = unlist(dist_to_tgts) %>% mean(),
         min_tgt_dist = unlist(dist_to_tgts) %>% min(),
         min_tgt_dist_no0 = unlist(dist_to_tgts) %>% Filter(gt0,.) %>% min())
df2_fc7 %>% 
  ggplot(aes(x = type, y = mean_tgt_dist)) + 
  stat_summary(fun.data = "mean_cl_boot", size = 1.5)

df2_fc7 %>% 
  ggplot(aes(x = type, y = min_tgt_dist)) + 
  stat_summary(fun.data = "mean_cl_boot", size = 1.5)

df2_fc7 %>% 
  ggplot(aes(x = type, y = min_tgt_dist_no0)) + 
  stat_summary(fun.data = "mean_cl_boot", size = 1.5)

```

#### Conv5

```{r}
df2_conv5_agg <- df2_conv5 %>% 
  rowwise() %>% 
  mutate(mean_tgt_dist = unlist(dist_to_tgts_conv5) %>% mean(),
         min_tgt_dist = unlist(dist_to_tgts_conv5) %>% min(),
         min_tgt_dist_no0 = unlist(dist_to_tgts_conv5) %>% Filter(gt0,.) %>% min())

df2_conv5_agg %>% 
  ggplot(aes(x = type, y = mean_tgt_dist)) + 
  stat_summary(fun.data = "mean_cl_boot", size = 1.5)

df2_conv5_agg %>% 
  ggplot(aes(x = type, y = min_tgt_dist)) + 
  stat_summary(fun.data = "mean_cl_boot", size = 1.5)

df2_conv5_agg %>% 
  ggplot(aes(x = type, y = min_tgt_dist_no0)) + 
  stat_summary(fun.data = "mean_cl_boot", size = 1.5)
```

#### Conv3

```{r}
df2_conv3_agg <- df2_conv3 %>% 
  rowwise() %>% 
  mutate(mean_tgt_dist = unlist(dist_to_tgts_conv3) %>% mean(),
         min_tgt_dist = unlist(dist_to_tgts_conv3) %>% min(),
         min_tgt_dist_no0 = unlist(dist_to_tgts_conv3) %>% Filter(gt0,.) %>% min())

df2_conv3_agg %>% 
  ggplot(aes(x = type, y = mean_tgt_dist)) + 
  stat_summary(fun.data = "mean_cl_boot", size = 1.5)

df2_conv3_agg %>% 
  ggplot(aes(x = type, y = min_tgt_dist)) + 
  stat_summary(fun.data = "mean_cl_boot", size = 1.5)

df2_conv3_agg %>% 
  ggplot(aes(x = type, y = min_tgt_dist_no0)) + 
  stat_summary(fun.data = "mean_cl_boot", size = 1.5)

```

## Load coords conv5

```{r}
id_subj <- df_participants %>% pull(subject_id) %>% unique()
coord_pth <- here("data/exp_DRM1/img_coords")

fs <- list.files(coord_pth,pattern = "*_conv5.rds", full.names = T)

for (i in 1:length(fs)) {
  fs1 <- fs[i]
  subj <- readRDS(fs1)

}

```