### 2020-04-24 reports/appendix.Rmd

---
title: "Appendix"
author: "Filip Děchtěrenko"
date: "24 4 2020"
output: html_document
---

```{r setup}
set.seed(190425)
library(tidyverse); theme_set(theme_classic(16))
library(lme4)
library(here)

source(here("R","utils.R"))

plots_dir <- here("plots", "exp_DRM1")

if(!dir.exists(plots_dir)) { dir.create(plots_dir) }

gt0 <- function(x) {x >0}

df2 <- df %>% pivot_longer(cols = starts_with("dist_to_tgts"), names_to = "layer", values_to = "distance_to_targets")
df2_dist_tgt <- df %>% pivot_longer(cols = starts_with("avg_dist_between_tgts"), names_to = "layer", values_to = "avg_dist_between_tgts")

df2 <- df2 %>% left_join(df2_dist_tgt %>% select(subject_id,trial_id), by = c("subject_id", "trial_id"))

df2_tgt <- df2 %>% 
  rowwise() %>% 
  mutate(mean_tgt_dist = unlist(distance_to_targets) %>% mean(),
         min_tgt_dist = unlist(distance_to_targets) %>% min(),
         min_tgt_dist_no0 = unlist(distance_to_targets) %>% Filter(gt0,.) %>% min()) %>% 
  ungroup() %>% 
  distinct()

divide_into_groups_tgt_dist_mean <- function(dfxx) {

  dfxx_notgt <- dfxx %>% filter(type != "target") %>% mutate(model_type = NA)

  rnk <- rank(dfxx_notgt$mean_tgt_dist)

  dfxx_notgt$model_type[rnk <= 5] <- "close_tgt"
  dfxx_notgt$model_type[rnk > 5]  <- "far_tgt"

  dfxx %>% 
    left_join(dfxx_notgt %>% select(img_name,model_type), by="img_name") %>% 
    mutate(model_type = if_else(type == "target","target",model_type))
}
divide_into_groups_tgt_dist_min_no0 <- function(dfxx) {

  dfxx_notgt <- dfxx %>% filter(type != "target") %>% mutate(model_type = NA)

  rnk <- rank(dfxx_notgt$min_tgt_dist_no0)

  dfxx_notgt$model_type[rnk <= 5] <- "close_tgt"
  dfxx_notgt$model_type[rnk > 5]  <- "far_tgt"

  dfxx %>% 
    left_join(dfxx_notgt %>% select(img_name,model_type), by="img_name") %>% 
    mutate(model_type = if_else(type == "target","target",model_type))
}

```

```{r}
df2_tgt_fc7 <- df2_tgt %>% 
  filter(layer == "dist_to_tgts")

df2_tgt_conv3 <- df2_tgt %>% 
  filter(layer == "dist_to_tgts_conv3")

df2_tgt_conv5 <- df2_tgt %>% 
  filter(layer == "dist_to_tgts_conv5")

```

# Conv5

## Divide distractors by distance to targets - mean, conv5

```{r}

df2_tgt_modeldist_conv5_mean <- df2_tgt_conv5 %>% 
  group_by(subject_id,grp_order) %>% 
  do(divide_into_groups_tgt_dist_mean(.))

df2_tgt_modeldist_conv5_mean %>% xtabs(~type+model_type,.) %>% prop.table(1)
df2_tgt_modeldist_conv5_mean %>% filter(type!="target") %>% xtabs(~type+model_type,.) %>% psych::phi()
df2_tgt_modeldist_conv5_mean %>% filter(type != "target") %>% 
  group_by(model_type) %>% 
  summarize(correct = mean(correct))

df2_tgt_modeldist_conv5_mean %>% filter(type != "target") %>% 
  group_by(type) %>% 
  summarize(correct = mean(correct))

```

## Divide distractors by distance to targets - min_no0, conv5

```{r}

df2_tgt_modeldist_conv5_min_no0 <- df2_tgt_conv5 %>% 
  group_by(subject_id,grp_order) %>% 
  do(divide_into_groups_tgt_dist_min_no0(.))

df2_tgt_modeldist_conv5_min_no0 %>% xtabs(~type+model_type,.) %>% prop.table(1)
df2_tgt_modeldist_conv5_min_no0 %>% filter(type!="target") %>% xtabs(~type+model_type,.) %>% psych::phi()
df2_tgt_modeldist_conv5_min_no0 %>% filter(type != "target") %>% 
  group_by(model_type) %>% 
  summarize(correct = mean(correct))

df2_tgt_modeldist_conv5_min_no0 %>% filter(type != "target") %>% 
  group_by(type) %>% 
  summarize(correct = mean(correct))

```

# Conv3

## Divide distractors by distance to targets - mean, conv3

```{r}

df2_tgt_modeldist_conv3_mean <- df2_tgt_conv3 %>% 
  group_by(subject_id,grp_order) %>% 
  do(divide_into_groups_tgt_dist_mean(.))

df2_tgt_modeldist_conv3_mean %>% xtabs(~type+model_type,.) %>% prop.table(1)
df2_tgt_modeldist_conv3_mean %>% filter(type!="target") %>% xtabs(~type+model_type,.) %>% psych::phi()
df2_tgt_modeldist_conv3_mean %>% filter(type != "target") %>% 
  group_by(model_type) %>% 
  summarize(correct = mean(correct))

df2_tgt_modeldist_conv3_mean %>% filter(type != "target") %>% 
  group_by(type) %>% 
  summarize(correct = mean(correct))

```

## Divide distractors by distance to targets - min_no0, conv3

```{r}

df2_tgt_modeldist_conv3_min_no0 <- df2_tgt_conv3 %>% 
  group_by(subject_id,grp_order) %>% 
  do(divide_into_groups_tgt_dist_min_no0(.))

df2_tgt_modeldist_conv3_min_no0 %>% xtabs(~type+model_type,.) %>% prop.table(1)
df2_tgt_modeldist_conv3_min_no0 %>% filter(type!="target") %>% xtabs(~type+model_type,.) %>% psych::phi()
df2_tgt_modeldist_conv3_min_no0 %>% filter(type != "target") %>% 
  group_by(model_type) %>% 
  summarize(correct = mean(correct))

df2_tgt_modeldist_conv3_min_no0 %>% filter(type != "target") %>% 
  group_by(type) %>% 
  summarize(correct = mean(correct))

```

# Figures

```{r}
df2_tgt_agg1 <- df2_tgt %>% 
    mutate(layer = recode(layer, dist_to_tgts = "fc7", dist_to_tgts_conv3 = "conv3", dist_to_tgts_conv5 = "conv5")) %>% 
    mutate(layer = factor(layer, levels = c("conv3", "conv5", "fc7")))

df2_tgt_agg1_ci <- df2_tgt_agg1 %>% 
  group_by(layer,type) %>% 
    summarize(m=Hmisc::smean.cl.boot(mean_tgt_dist)[1],lower =Hmisc::smean.cl.boot(mean_tgt_dist)[2], upper = Hmisc::smean.cl.boot(mean_tgt_dist)[3])

p <- df2_tgt_agg1_ci %>% 
  group_by(layer) %>% 
  mutate(upper = upper/upper[type=="target"], lower = lower/lower[type=="target"], m = m/m[type=="target"]) %>% 
  ggplot(aes(x=layer, y=m,col = type, group = type)) + 
  geom_point()+geom_line() +
  geom_linerange(aes(ymin=lower,ymax=upper))+
  ylab("Normalized average distance to targets") + 
  theme(aspect.ratio = 1)
p
ggsave(file.path(plots_dir, "XXAppDistances_mean.svg"), p, width = 6, height = 6)

df2_tgt_agg2_ci <- df2_tgt_agg1 %>% 
  group_by(layer,type) %>% 
  summarize(m=Hmisc::smean.cl.boot(min_tgt_dist_no0)[1],lower =Hmisc::smean.cl.boot(min_tgt_dist_no0)[2], upper = Hmisc::smean.cl.boot(min_tgt_dist_no0)[3])

p <- df2_tgt_agg2_ci %>% 
  group_by(layer) %>% 
  mutate(upper = upper/upper[type=="target"], lower = lower/lower[type=="target"], m = m/m[type=="target"]) %>% 
  ggplot(aes(x=layer, y=m,col = type, group = type)) + 
  geom_point()+geom_line() +
  geom_linerange(aes(ymin=lower,ymax=upper))+
  ylab("Normalized minimum distance to targets") + 
  theme(aspect.ratio = 1)
p
ggsave(file.path(plots_dir, "XXAppDistances_min.svg"), p, width = 6, height = 6)
```